---
title: "Random Forest Model File"
author: "Jingbin Cao, jc5514"
date: "3/13/2021"
output: pdf_document
---
# Getting all packages
```{r message=FALSE, warning=FALSE}
if(!require("EBImage")){
 install.packages("BiocManager")
 BiocManager::install("EBImage")
}
if(!require("R.matlab")){
  install.packages("R.matlab")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("dplyr")){
  install.packages("dplyr")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("ggplot2")){
  install.packages("ggplot2")
}

if(!require("caret")){
  install.packages("caret")
}

if(!require("glmnet")){
  install.packages("glmnet")
}

if(!require("WeightedROC")){
  install.packages("WeightedROC")
}

if(!require("gbm")){
  install.packages("gbm")
}

if(!require("DMwR")){
  install.packages("DMwR")
}
if(!require("R.matlab")){
  install.packages("R.matlab")
}
if(!require("randomForest")){
 install.packages("randomForest")
}
if(!require("xgboost")){
 install.packages("xgboost")
}
if(!require("tibble")){
 install.packages("tibble")
}
if(!require("ROSE")){
 install.packages("ROSE")
}
if(!require("ggplot2")){
 install.packages("ggplot2")
}
if(!require("tidyverse")){
 install.packages("tidyverse")
}

if(!require("AUC")){
 install.packages("AUC")
}
if(!require("e1071")){
 install.packages("e1071")
}
if(!require("OpenImageR")){
 install.packages("OpenImageR")
}
if(!require("caTools")){
  install.packages("caTools")
}

library(R.matlab)
library(readxl)
library(dplyr)
library(EBImage)
library(ggplot2)
library(caret)
library(glmnet)
library(WeightedROC)
library(gbm)
library(DMwR)
library(OpenImageR)
library(AUC)
library(e1071)
library(randomForest)
library(xgboost)
library(tibble)
library(ROSE)
library(ggplot2)
library(tidyverse)
library(AUC)
library(e1071)
library(caTools)
library(R.matlab)
```

# All controls you need to adjust
```{r}
run.cv <- FALSE # run cross-validation on the training set
sample.reweight <- FALSE # run sample reweighting in model training
K <- 5  # number of CV folds
gbm.numtrees <- 1000 #number of trees to use in gbm
run.feature.train <- FALSE # process features for training set
run.test <- TRUE # run evaluation on an independent test set
run.feature.test <- FALSE # process features for test set
run.poly.feature <- TRUE # process poly features
run.add.poly.feature <- TRUE # and poly features to dist matrix
run.gbm <- F
needs.balanced <- TRUE # balance data for model fitting
model.selection <- TRUE # perform model selection on svm models
# Random Forest Model with new feature
run.rf <- FALSE
train.rf <- FALSE # Train Random Forest Model
test.rf <- TRUE # Test Random Forest Model
# Random Forest Model with old feature
run.rf.old.feature <- TRUE
train.rf.old.feature <- TRUE
test.rf.old.feature <- TRUE
```

# Setting Data
## Getting raw data ready (Same as main.rmd)
```{r}
set.seed(2020)
setwd("../doc")
# All directories for training images:
train_dir <- "../data/train_set/" # This will be modified for different data sets.
train_image_dir <- paste(train_dir, "images/", sep="")
train_pt_dir <- paste(train_dir,  "points/", sep="")
train_label_path <- paste(train_dir, "label.csv", sep="")
```

## Import raw data
```{r}
#train-test split
info <- read.csv(train_label_path)
n <- nrow(info) #get number of rows from csv
n_train <- round(n*(4/5), 0) #use 4/5 amount of data for training
train_idx <- sample(info$Index, n_train, replace = F) #grab indexs used for training
test_idx <- setdiff(info$Index, train_idx) # get indexs not used for training

# Below are from starter code
n_files <- length(list.files(train_image_dir,'*jpg'))
```
## Read Fiducial Points
```{r}
readMat.matrix <- function(index){
     return(round(readMat(paste0(train_pt_dir, sprintf("%04d", index),".mat"))[[1]],0))
}

#load fiducial points
fiducial_pt_list <- lapply(1:n_files, readMat.matrix)
save(fiducial_pt_list, file="../output/fiducial_pt_list.RData")
```
## Getting all features ready
```{r}
source("../lib/feature.R")
tm_feature_train <- NA
if(run.feature.train){
  tm_feature_train <- system.time(dat_train<-feature(fiducial_pt_list,train_idx, run.poly.feature, run.add.poly.feature))
  save(dat_train, file="../output/feature_train_RF.RData")
}else{
  load(file="../output/feature_train_RF.RData")
}
tm_feature_test <- NA
if(run.feature.test){
  tm_feature_test <- system.time(dat_test <- feature(fiducial_pt_list, test_idx, run.poly.feature, run.add.poly.feature))
  save(dat_test, file="../output/feature_test_RF.RData")
}else{
  load(file="../output/feature_test_RF.RData")
}
# transfer label column from factor to numeric

dat_train$label <- as.numeric(dat_train$label)
dat_test$label <- as.numeric(dat_test$label)
```
# Advanced Model 2: Random Forest (RF) Model with new features

The second advanced model is random forest. Here we use the datasets that are extracted by new feature functions. We used two models trained by both imbalanced and balanced dataset. We used ROSE function to balance bothh training and testing data. For tuning the model, we set mtry = 308, tree number = 500, and node size = 10 for the RF model using balanced data, and we set mtry = 308, tree number = 1500, and node size = 30 for the RF model using imbalanced data. The evaluation of the model is shown at the end of this section.  

The tuning part is in a separate file named "appendix_tune_rf.rmd" in doc folder. Please feel free to check that to see the tuning process.  

We also trained random forest model with the datasets that are extracted by old feature functions. That is in part 6. Thank you for reading!
```{r}
if(run.rf){
#balanced data
dat_train_balanced_rose<-ROSE(label~., dat_train,seed=2020)$data
dat_test_balanced_rose <- ROSE(label~., dat_test, seed=2020)$data


## Training RF

source("../lib/random_forest.R")
if(train.rf){
  # Train RF by balanced data
  time.rf.train.final.balanced <- system.time(random_forest_fit_final_balanced <- random_forest_train(dat_train_balanced_rose, mtry = 308, tree_number = 500, node_size = 10))
  save(random_forest_fit_final_balanced, file = "../output/rf_train_final_balanced.RData")
  save(time.rf.train.final.balanced, file = "../output/rf_train_final_time_balanced.RData")
  # Train RF by imbalanced data
  time.rf.train.final.imbalanced <- system.time(random_forest_fit_final_imbalanced <- random_forest_train(dat_train, mtry = 308, tree_number = 1500, node_size = 30))
  save(time.rf.train.final.imbalanced, file = "../output/rf_train_final_time_imbalanced.RData")
  save(random_forest_fit_final_imbalanced, file = "../output/rf_train_final_imbalanced.RData")
}else{
  load("../output/rf_train_final_balanced.RData")
  load("../output/rf_train_final_time_balanced.RData")
  load("../output/rf_train_final_time_imbalanced.RData")
  load("../output/rf_train_final_imbalanced.RData")
}
# Evaluation:
# Balanced:
if(test.rf){
time.rf.test.final.balanced <- system.time(rf_predicted_balanced <- as.numeric(as.vector(random_forest_test(random_forest_fit_final_balanced, dat_test_balanced_rose))))
rf_accuracy_balanced <- mean(round(rf_predicted_balanced == dat_test_balanced_rose$label))
tpr.fpr.balanced <- WeightedROC(as.numeric(rf_predicted_balanced),dat_test_balanced_rose$label)
rf_AUC_balanced <- WeightedAUC(tpr.fpr.balanced)

cat("AUC for tuned Random Forest(balanced): ", rf_AUC_balanced,".\n")
cat("Accuracy for tuned Random Forest(balanced)", rf_accuracy_balanced*100,"%.\n")
cat("Training time for tuned Random Forest: ", time.rf.train.final.balanced[1], "s.\n")
cat("Testing time for tuned Random Forest: ", time.rf.test.final.balanced[1], "s.\n")
cat("   ","\n")
# Imbalanced:
time.rf.test.final.imbalanced <- system.time(rf_predicted_imbalanced <- as.numeric(as.vector(random_forest_test(random_forest_fit_final_imbalanced, dat_test))))
rf_accuracy_imbalanced <- mean(round(rf_predicted_imbalanced == dat_test$label))
tpr.fpr.imbalanced <- WeightedROC(as.numeric(rf_predicted_imbalanced),dat_test$label)
rf_AUC_imbalanced <- WeightedAUC(tpr.fpr.imbalanced)

cat("AUC for tuned Random Forest(imbalanced): ", rf_AUC_imbalanced,".\n")
cat("Accuracy for tuned Random Forest(imbalanced)", rf_accuracy_imbalanced*100,"%.\n")
cat("Training time for tuned Random Forest: ", time.rf.train.final.imbalanced[1], "s.\n")
cat("Testing time for tuned Random Forest: ", time.rf.test.final.imbalanced[1], "s.\n")
}
}
```
We think RF model do not need Cross-Validation. Here is a snippet from Breiman's official documentation:
In random forests, there is no need for cross-validation or a separate test set to get an unbiased estimate of the test set error. It is estimated internally, during the run, as follows: Each tree is constructed using a different bootstrap sample from the original data. About one-third of the cases are left out of the bootstrap sample and not used in the construction of the kth tree. Put each case left out in the construction of the kth tree down the kth tree to get a classification. In this way, a test set classification is obtained for each case in about one-third of the trees. At the end of the run, take j to be the class that got most of the votes every time case n was oob. The proportion of times that j is not equal to the true class of n averaged over all cases is the oob error estimate. This has proven to be unbiased in many tests.


#  Advanced Model 6: Random Forest (RF) Model with old feature:  
The sixth advanced model is random forest. Here we use the datasets that are extracted by old feature functions. We used two models trained by both imbalanced and balanced dataset. We used ROSE function to balance bothh training and testing data. For tuning the model, we set mtry = 308, tree number = 500, and node size = 10 for the RF model using balanced data, and we set mtry = 308, tree number = 1500, and node size = 30 for the RF model using imbalanced data. The evaluation of the model is shown at the end of this section, and we will compare this model with the RF model trained by new features.
```{r}
if(run.rf.old.feature){
source("../lib/random_forest_old_feature.R")
old_feature <- function(input_list = fiducial_pt_list, index){
  old_pairwise_dist <- function(vec){
    return(as.vector(dist(vec)))
  }
  old_pairwise_dist_result <-function(mat){
    return(as.vector(apply(mat, 2, old_pairwise_dist))) 
  }
  old_pairwise_dist_feature <- t(sapply(input_list[index], old_pairwise_dist_result))
  dim(old_pairwise_dist_feature) 
  old_pairwise_data <- cbind(old_pairwise_dist_feature, info$label[index])
  colnames(old_pairwise_data) <- c(paste("feature", 1:(ncol(old_pairwise_data)-1), sep = ""), "label")
  old_pairwise_data <- as.data.frame(old_pairwise_data)
  old_pairwise_data$label <- as.factor(old_pairwise_data$label)
  return(feature_df = old_pairwise_data)
}

old_tm_feature_train <- NA
if(run.feature.train){
  old_tm_feature_train <- system.time(old_dat_train <- old_feature(fiducial_pt_list, train_idx))
  save(old_dat_train, file="../output/feature_train_old.RData")
}else{
  load(file="../output/feature_train_old.RData")
}
old_tm_feature_test <- NA
if(run.feature.test){
  old_tm_feature_test <- system.time(old_dat_test <- old_feature(fiducial_pt_list, test_idx))
  save(old_dat_test, file="../output/feature_test_old.RData")
}else{
  load(file="../output/feature_test_old.RData")
}
# transfer label column from factor to numeric

old_dat_train$label <- as.numeric(old_dat_train$label)
old_dat_test$label <- as.numeric(old_dat_test$label)

# Balance data
old_dat_train_balanced_rose <- ROSE(label~., old_dat_train,seed=2020)$data
old_dat_test_balanced_rose <- ROSE(label~., old_dat_test, seed=2020)$data


# Train Model
if(train.rf.old.feature){
  # Balanced
  old_time.rf.train.final.balanced <- system.time(old_random_forest_fit_final_balanced <- old_random_forest_train(old_dat_train_balanced_rose, mtry = 308, tree_number = 500, node_size = 10))
  save(old_random_forest_fit_final_balanced, file = "../output/rf_train_final_balanced_old_feature.RData")
  save(old_time.rf.train.final.balanced, file = "../output/rf_train_final_time_balanced_old_feature.RData")
  # Imbalanced
  old_time.rf.train.final.imbalanced <- system.time(old_random_forest_fit_final_imbalanced <- old_random_forest_train(old_dat_train, mtry = 308, tree_number = 1500, node_size = 30))
  save(old_time.rf.train.final.imbalanced, file = "../output/rf_train_final_time_imbalanced_old_feature.RData")
  save(old_random_forest_fit_final_imbalanced, file = "../output/rf_train_final_imbalanced_old_feature.RData")
}else{
  load("../output/rf_train_final_balanced_old_feature.RData")
  load("../output/rf_train_final_time_balanced_old_feature.RData")
  load("../output/rf_train_final_time_imbalanced_old_feature.RData")
  load("../output/rf_train_final_imbalanced_old_feature.RData")
}


# Evaluate Model
# Balanced:
if(test.rf.old.feature){
old_time.rf.test.final.balanced <- system.time(rf_predicted_balanced <- as.numeric(as.vector(old_random_forest_test(old_random_forest_fit_final_balanced, old_dat_test_balanced_rose))))
old_rf_accuracy_balanced <- mean(round(rf_predicted_balanced == old_dat_test_balanced_rose$label))
old_tpr.fpr.balanced <- WeightedROC(as.numeric(rf_predicted_balanced),old_dat_test_balanced_rose$label)
old_rf_AUC_balanced <- WeightedAUC(old_tpr.fpr.balanced)

cat("AUC(balanced) for Random Forest with old feature: ", old_rf_AUC_balanced,".\n")
cat("Accuracy(balanced) for Random Forest with old feature", old_rf_accuracy_balanced*100,"%.\n")
cat("Training time (balanced) for Random Forest with old feature: ", old_time.rf.train.final.balanced[1], "s.\n")
cat("Testing time (balanced) for Random Forest with old feature: ", old_time.rf.test.final.balanced[1], "s.\n")
cat("   ","\n")

# Imbalanced:
old_time.rf.test.final.imbalanced <- system.time(rf_predicted_imbalanced <- as.numeric(as.vector(old_random_forest_test(old_random_forest_fit_final_imbalanced, old_dat_test))))
old_rf_accuracy_imbalanced <- mean(round(rf_predicted_imbalanced == old_dat_test$label))
old_tpr.fpr.imbalanced <- WeightedROC(as.numeric(rf_predicted_imbalanced),old_dat_test$label)
old_rf_AUC_imbalanced <- WeightedAUC(old_tpr.fpr.imbalanced)

cat("AUC(imbalanced) for Random Forest with old feature: ", old_rf_AUC_imbalanced,".\n")
cat("Accuracy(imbalanced) for Random Forest with old feature", old_rf_accuracy_imbalanced*100,"%.\n")
cat("Training time (imbalanced) for Random Forest with old feature: ", old_time.rf.train.final.imbalanced[1], "s.\n")
cat("Testing time (imbalanced) for Random Forest with old feature: ", old_time.rf.test.final.imbalanced[1], "s.\n")
}
}
```
